---
title: "Finding The Next Big Hit Through Danceability"
subtitle: "Report"
format: html
editor: visual
execute:
  echo: true
---

## Introduction and data

Our research project will center around the spotifyr, an open source music dataset provided by Spotify. The dataset contains around 5,000 songs from six major music categories (EDM, Latin, Pop, R&B, Rap, and Rock). For each song there contains general information such as the songs album, artist, release data, but also more interesting statistics on danceability, energy, and popularity. The Spotify Web API was used to gain information about the songs. Spotify has twelve different features that are associated with songs. These are included within the dataset as variables.

We considered different ethical questions surrounding data collection. Specifically, we explored how songs were chosen from Spotify and acquisition bias; however, do not believe there are significant ethical concerns related to this data and data collection.

Within the music industry there is always a desire to be able to predict the next big song. Different factors will always play into whether or not a song becomes popular. We wish to explore how danceability, in particular, affects the popularity of songs. To dig a bit deeper, we want to see if the impact of danceability on popularity changes depending on which genre we are exploring. We are hoping that by exploring this correlation we can create a predictive model for popularity based on danceability. Based on the correlations we find by genre this model may be more accurate for certain genres more so than others. 

We hypothesize that there will be a positive correlation between danceability and popularity, but that correlation will differ between genres. We expect the Latin genre to have the greatest correlation. Additionally, we hypothesis that the tempo and the energy of the song (when taken into consideration with the danceability rating) will better be able to predict track popularity rating.

## Methodology

We plan to answer this research question through a linear regression analysis looking at if the danceability of a song can be predictive for its popularity. First, we will look to see at the general relationship between danceability rating and track popularity rating of songs in the data. Then, we will look to into the level of this relationship in terms of different types of music by assessing genre as variable. In addition to these basic visualizations, we will also calculate a number of summary statistics that will further assist in other analyses and allow us to better understand the data. This includes calculation of the mean danceability rating and popularity ratings specifically to each genre.

Additionally, we will conduct a regression analysis to assess how other variables influence the relationship between danceability and popularity. The approriate type of regression for our data would be a multiple linear regression that would take into consideration multiple explanatory variables to predict track popularity. The specific quantitative variables we want to consider are energy and tempo because of their relation to danceability in this regression analysis as similarly considered by a similar project by Ochi et al. (n.d.). We will assess these types of regressions with both additive and interaction models to see which type of model best fits the data. This relationship will also be analyzed for individual genres as well.

## Data

```{r}
#| label: load-libraries
#| message: false
#| warning: false

library(tidyverse)
```

```{r}
#| label: load-data
#| message: false
#| warning: false

spotify_songs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-21/spotify_songs.csv')

head(spotify_songs)
```

## Analysis

```{r}
#| label: popularity-vs-danceability-general

spotify_songs |>
  ggplot(
    aes(
      x = danceability,
      y = track_popularity
    )
  ) +
  geom_point(color = "steelblue") +
  geom_smooth(se = FALSE) +
  labs(
    x = "Danceability Rating",
    y = "Track Popularity Rating",
    title = "How does danceability affect the popularity of a song?"
  )
```
