---
title: "Finding The Next Big Hit Through Danceability"
subtitle: "Report"
format: html
editor: visual
execute:
  echo: true
---

## Introduction and data

Our research project will center around the spotifyr, an open source music dataset provided by Spotify. The dataset contains around 5,000 songs from six major music categories (EDM, Latin, Pop, R&B, Rap, and Rock). For each song there contains general information such as the songs album, artist, release data, but also more interesting statistics on danceability, energy, and popularity. The Spotify Web API was used to gain information about the songs. Spotify has twelve different features that are associated with songs. These are included within the dataset as variables.

We considered different ethical questions surrounding data collection. Specifically, we explored how songs were chosen from Spotify and acquisition bias; however, do not believe there are significant ethical concerns related to this data and data collection.

Within the music industry there is always a desire to be able to predict the next big song. Different factors will always play into whether or not a song becomes popular. We wish to explore how danceability, in particular, affects the popularity of songs. To dig a bit deeper, we want to see if the impact of danceability on popularity changes depending on which genre we are exploring. We are hoping that by exploring this correlation we can create a predictive model for popularity based on danceability. Based on the correlations we find by genre this model may be more accurate for certain genres more so than others. 

We hypothesize that there will be a positive correlation between danceability and popularity, but that correlation will differ between genres. We expect the Latin genre to have the greatest correlation. 

## Methodology

We plan to answer this research question through a linear regression analysis looking at if the danceability of a song can be predictive for its popularity. This regression analysis will likely consider the energy and tempo variables in the regression analysis. This analysis will be faceted across genres as well.

## Data

```{r}
#| label: load-libraries
#| message: false
#| warning: false

library(tidyverse)
```

```{r}
#| label: load-data
#| message: false
#| warning: false

spotify_songs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-21/spotify_songs.csv')

head(spotify_songs)
```

## Analysis

```{r}
#| label: popularity-vs-danceability-general

spotify_songs |>
  ggplot(
    aes(
      x = danceability,
      y = track_popularity
    )
  ) +
  geom_point(color = "steelblue") +
  geom_smooth(se = FALSE) +
  labs(
    x = "Danceability Rating",
    y = "Track Popularity Rating",
    title = "How does danceability affect the popularity of a song?"
  )
```
